Handling String-Number Conversion in String Constraint Solving

### 2019-11-07 draft.tex
add new paper structure

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview} \label{section:overview}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
For proving UNSAT, our strategy first over-approximates string-number conversion function $n=\its{x}$ as an uninterpreted function with an additional axiom $n\leq 0 \rightarrow \sti{\its{n}}=n$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Symbolic Flat Automata} \label{section:sfa}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\st}[2]{q_{#1}^{#2}}
\newcommand{\sym}[2]{v_{#1}^{#2}}
%We use $\cvar$ to denote a finite set of \emph{character variables}  ranging over $\Sigma_\epsilon$. 
%Sometimes we use index to denote different set of variables over the same range, e.g., $\cvarone$ and $\cvartwo$. 
%We use $v$ and its variants for character variables. We call a string over $\cvar$ (character variables) a \emph{symbolic string}.
%An \emph{interpretation $I$ over $\cvar$} is a mapping from $\cvar$ to $\Sigma_\epsilon$. For a symbolic string $x= v_1v_2\cdots v_k$ over $\cvar$, $I(x)$ is defined as $I(v_1)\cdot I(v_2)\cdot \ldots \cdot I(v_k)$.

@@

%\section{Handling Equality and Disequality Constraint Symbolically} \label{section:eq}
\section{Flat Under-approximation for Equalities and Regular Constraints} \label{section:eq}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Handling String-Integer Conversion} \label{section:s2i}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\todo{talk about the design principle}
In order to handle the under-approximation of string-integer conversion efficiently, we propose to restrict the domain of $x$ to a SFA in Figure~\ref{fig:sfa_its}.

The SFA has a self-loop transition labeled $v_0$ at $q_0$ and we always add a constraint $v_0=0$ for such SFA. With this transition, we can ensure that the under-approximation handles all numbers with at most $m$ digits. Notice that even for a bounded integer, the corresponding string can be of unbounded length with arbitrarily many `$0$' characters at the front. For example, the formula $\sti{x}=10 \wedge |x|=5$ is satisfiable when $x=``00010"$.

@@

Such SFA handles bounded integers (e.g., 10-digits or 64-bits) precisely and under-approximate the possible values of unbounded integers. It works very efficiently because we just need to use linear constraints to characterize the relation between $x$ and $n$. 
Remember that the character variables can be assigned $\epsilon$. In order to handle such cases efficiently, we use an encoding that if $v_j = \epsilon = -1$ then $v_k = -1$ for all $k\in [j,m]$, i.e., we shift all the $\epsilon$-transitions to the least significant digits. More specifically, we first define two functions:

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview} \label{section:overview}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo{Some non-technical text to explain how the approach solves a toy problem}
For proving UNSAT, our strategy first over-approximates string-number conversion function $n=\its{x}$ as an uninterpreted function with an additional axiom $n\leq 0 \rightarrow \sti{\its{n}}=n$.

@@

\section{Under-approximation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Symbolic Flat Automata} \label{section:sfa}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\st}[2]{q_{#1}^{#2}}
\newcommand{\sym}[2]{v_{#1}^{#2}}
%We use $\cvar$ to denote a finite set of \emph{character variables}  ranging over $\Sigma_\epsilon$. 
%Sometimes we use index to denote different set of variables over the same range, e.g., $\cvarone$ and $\cvartwo$. 
%We use $v$ and its variants for character variables. We call a string over $\cvar$ (character variables) a \emph{symbolic string}.
%An \emph{interpretation $I$ over $\cvar$} is a mapping from $\cvar$ to $\Sigma_\epsilon$. For a symbolic string $x= v_1v_2\cdots v_k$ over $\cvar$, $I(x)$ is defined as $I(v_1)\cdot I(v_2)\cdot \ldots \cdot I(v_k)$.

@@

%\section{Handling Equality and Disequality Constraint Symbolically} \label{section:eq}
\subsection{Flat Under-approximation for Equalities and Regular Constraints} \label{section:eq}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Handling String-Integer Conversion} \label{section:s2i}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\paragraph{SFA that allows efficient string-integer conversion: } In the previous discussion, we know that it is unlikely to use arbitrary SFA and gain efficient solving procedure, due to the requirement to solve exponential integer constraints. The problem is that each loop in the SFA creates an exponential component. An immediate idea would be that maybe we can restrict the variable to some specific SFA that allow more efficient constraint solving. At the same time, we want to keep some good property, i.e., the search space covers all integers with at most $m$-digits.

Our first attempt is a SFA with no loop, i.e., a straight line structure (e.g., the SFA in Fig~\ref{fig:sfa_its} without the self-loop at $q_0$). The corresponding integer constraint does not have exponential components. However, it does not ensure the property that all numbers with at most $m$-digits are covered. Consider the example $\sti{x}=10 \wedge |x|=5$. The number $10$ has only $2$-digits, at the first glance, a straight-line SFA with two transitions, i.e., the SFA \scalebox{0.5}{\tikzset{state/.style={circle,draw=blue!50,fill=blue!20,
			thick,inner sep=0pt,minimum size=6mm}, initial text=$ $}
	\begin{tikzpicture} 
	\node[state,initial] (q0)  {$q_0$};
	\node[state] (q1) [right = 1cm of q0] {$q_1$};
	\node[state,accepting] (q2) [right = 1cm of q1] {$q_2$};

	\draw[->] (q0) edge node [below]{$v_1$} (q1) ;
	\draw[->] (q1) edge node [below]{$v_2$} (q2) ;
	\end{tikzpicture}}, should be sufficient for the domain restriction of $x$. If we do so, we will conclude that the formula is unsatisfiable under this domain restriction.

However, the formula is satisfiable when $x=``00010"$. The point is that even for a bounded integer, the corresponding string can be of unbounded length with arbitrarily many `$0$' characters at the front. In order to handle all number with at most $m$ digits, we propose to restrict the domain of $x$ to a SFA in Figure~\ref{fig:sfa_its}.  The SFA has a self-loop transition labeled $v_0$ at $q_0$ and we always add a constraint $v_0=0$ for such SFA. With this transition, we can ensure that the under-approximation handles all numbers with at most $m$ digits. Although this SFA still has a loop, but it will not create any exponential component because the loop only represents a sequence of `0' that occurs in the prefix of $x$. It will not affect the integer value of $n=\sti{x}$.

@@

Such SFA handles bounded integers (e.g., 10-digits or 64-bits) precisely and under-approximate the possible values of unbounded integers. It works very efficiently because we just need to use linear constraints to characterize the relation between $x$ and $n$. 
Remember that the character variables can be assigned $\epsilon$. In order to handle such cases efficiently, we use an encoding that if $v_j = \epsilon = -1$ then $v_k = -1$ for all $k\in [j,m]$, i.e., we shift all the $\epsilon$-transitions to the least significant digits. In order to explain this encoding, we first define two macro function:

More specifically, we first define two functions:
